---
title: "HW2"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```

## 'Problem 1' 

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or case_match function may be useful).

Import dataset and clean up column names 
```{r}
subway_df = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")

subway_df = janitor::clean_names(subway_df)
```


Retain specified columns and convert entry variable from char to log

```{r}
subway_df = 
  select(subway_df, line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, entrance_type, ada) |>
  mutate(entry = ifelse(entry == "NO", 0, 1))
```

This dataset contains variables about the NYC subway lines. Columns include line, station name, coordinates, routes, entry, vending, entrance_type, and ADA accessibility. It has 1868 rows and 
19 columns. To clean the data, I first used the janitor step to clean the column names to use snake case formatting. Then I selecting our columns of interst and changed the entry variable from character to logical. These data are not fully tidy because the routes contain both doubles and character variables. 

