---
title: "HW2"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
```

## 'Problem 1' 

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or case_match function may be useful).

Import dataset and clean up column names. Retain specified columns and convert entry variable from char to log. Convert route 8-11 to character variables to match route 1-7.
```{r}
subway_df = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
  select(line, station_name, station_latitude, station_longitude, route1, route2,route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, entrance_type, ada) |>
   mutate(
      route8 = as.character(route8),
      route9 = as.character(route9),
      route10 = as.character(route10),
      route11 = as.character(route11)) |>
  mutate(entry = ifelse(entry == "NO", FALSE, TRUE)) 
```

Apply pivot_longer to tidy up route columns.  

```{r}
subway_df <- subway_df |> 
  pivot_longer(
    cols = route1:route11,
    names_to = "route", 
    names_prefix = "route",
    values_to = "subway_line")
```

This dataset contains variables about the NYC subway lines. Columns include line, station name, coordinates, route,subway_route, entry, vending, entrance_type, and ADA accessibility. It has 4270 rows and 10 columns. To clean the data, I first used the janitor step to clean the column names to use snake case formatting. Then I selected our columns of interest and changed the entry variable from character to logical. Then I converted all the routes to character variables so I could apply pivot-longer to get route1-route11 in rows rather than columns. 


How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.

```{r}
subway_df$full_station_name <-
  paste(subway_df$line, subway_df$station_name)
```
There are `r nrow(distinct(subway_df, full_station_name))` distinct station names. 

How many stations are ADA compliant?
```{r}
sum(subway_df$ada, na.rm=TRUE)
```
There are `r nrow(sum(subway_df$ada, na.rm=TRUE))` stations


What proportion of station entrances / exits without vending allow entrance?
```{r}

```

### Problem 2 

Read and clean the Mr. Trash Wheel sheet:

specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
use reasonable variable names
omit rows that do not include dumpster-specific data
round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

```{r}
mr_trash_wheel = read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                            sheet = "Mr. Trash Wheel", range = "A2:N653") |>
  janitor::clean_names() |>
   drop_na(dumpster) |>
  mutate(
    sports_balls = as.integer(round(sports_balls))
    )
```

Use a similar process to import, clean, and organize the data for Professor Trash Wheel and Gwynnda.

```{r}
prof_trash_wheel = read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                            sheet = "Professor Trash Wheel", range = "A2:M121") |>
  janitor::clean_names() |>
   drop_na(dumpster) 

gwynnda_trash_wheel = read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                            sheet = "Gwynnda Trash Wheel", range = "A2:L266") |>
  janitor::clean_names() |>
   drop_na(dumpster) 
```


Combine this with the Mr. Trash Wheel dataset to produce a single tidy dataset. To keep track of which Trash Wheel is which, you may need to add an additional variable to both datasets before combining.

```{r}
mr_trash_wheel <- mr_trash_wheel |>
  mutate(wheel = "trash_wheel" ,
          year = as.double(year))

prof_trash_wheel<- prof_trash_wheel |>
  mutate(wheel = "prof_trash_wheel", 
          year = as.double(year))

gwynnda_trash_wheel  <- gwynnda_trash_wheel |>
  mutate(wheel = "gwynnda_wheel",
          year = as.double(year))

trash_wheel_tidy <-
  bind_rows(mr_trash_wheel, prof_trash_wheel, gwynnda_trash_wheel)
```


Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in the resulting dataset, and give examples of key variables. For available data, what was the total weight of trash collected by Professor Trash Wheel? What was the total number of cigarette butts collected by Gwynnda in June of 2022?


These data ...
The number of observations in this dataset are 1033 and key variables are wheel, dumpster, month, year, date, and volume_cubic_yards.


The total weight of trash collected by Professor Trash Wheel is 

`r sum(pull(filter(trash_wheel_tidy, wheel == "prof_trash_wheel"), weight_tons),na.rm = TRUE)` tons total weight of trash. 



The total number of cigarette butts collected by Gwynnda in June of 2022 

```{r}
trash_wheel_tidy_butts <- trash_wheel_tidy |>
  mutate(year = as.numeric(year)) |>
  filter( month == "June", year == 2022)
```

The total number of cigarette butts collected by Gwynnda in June of 2022 `r sum(pull(filter(trash_wheel_tidy_butts, wheel == "gwynnda_wheel"), cigarette_butts), na.rm = TRUE) ` cigarette butts.


### Problem 3

In the first part of this problem, your goal is to create a single, well-organized dataset with all the information contained in these data files. To that end: import, clean, tidy, and otherwise wrangle each of these datasets; check for completeness and correctness across datasets (e.g. by viewing individual datasets and using anti_join); merge to create a single, final dataset; and organize this so that variables and observations are in meaningful orders. Export the result as a CSV in the directory containing the original datasets.

Using janitor::clean_names() to clean all the column names.   

```{r}
bakers_df = read_csv("data/bakers.csv") |>
  janitor::clean_names() 
  
bakes_df = read_csv("data/bakes.csv",
  na = c("NA", ".", "", "UNKNOWN","Unknown", "N/A")) |>
  janitor::clean_names() 
  
results_df = read_csv("data/results.csv", skip = 2) |>
  janitor::clean_names() 
  
viewers_df = read_csv("data/viewers.csv") |>
  janitor::clean_names() 
```

Baker_name in bakers_f is first and last name whereas all other df has column baker with just first name. Separate this column into first and last name. 

Results_df first two rows are not useful, skip them so the third row becomes our columns. 

Bakes_df = 

We notice that in each of the four datasets there is series so we can join by that variable and want to make sure it is organized and the same across all four datasets. 

clean bakers_df
```{r}
bakers_df <- bakers_df |>
  separate(baker_name, into = c("baker", "baker_last_name"), sep = " ")

```

clean viewers_df
```{r}
viewers_df <- viewers_df |>
  pivot_longer(
    cols = series_1:series_10,
    names_to = "series_number",
    names_prefix = "series_",
    values_to = "series"
    )
```

## Check for completeness and correctness 
```{r}
missing_bakers_df <- bakers_df |>
  anti_join(bakes_df, by = c("baker", "series"))

```
There are `r nrow(missing_bakers_df)` - 1(row for column names) = 25 missing bakers that do not
have the same information in bakes_df

```{r}
  anti_join(bakers_df, bakes_df)
```
There are 25 missing bakers that do not
have the same information in bakes_df

```{r}
anti_join(bakers_df, results_df)
```
Jo Wheatley is missing, it appears that in this is a nickname and her full name is Joanne. 
Let's fix this. 

```{r}
results_df <- results_df |> 
  mutate(baker = case_match(
         baker, "Joanne" ~ "Jo")
  )
```


### Merge to create a single, final dataset; and organize this so that variables and observations are in meaningful orders.

Left join to merge datasets
```{r}
bakers_bakes_merge = 
  left_join(bakers_df, bakes_df, by = c("series", "baker")) 
```

Another merge
```{r}
results_bakersbakes_merge = 
  left_join(bakers_bakes_merge, results_df, by = c("series", "baker"))
```

Final merge

First rename columns in viewers_df
```{r}
viewers_df <- viewers_df |>
  rename(viewers = series,
          series = series_number) |>
  mutate(series = as.double(series))
```

```{r}
full_merge = 
  left_join(results_bakersbakes_merge, viewers_df, by = c("series")) 

write_csv(full_merge, "data/full_merge.csv")
```

Describe your data cleaning process, including any questions you have or choices you made. Briefly discuss the final dataset.

This process was difficult. I still have a hard time understanding pivot moves and why they are useful. 


Create a reader-friendly table showing the star baker or winner of each episode in Seasons 5 through 10. Comment on this table – were there any predictable overall winners? Any surprises?
```{r}
star_baker_or_winner <- full_merge |>
  filter(
          result %in% c("STAR BAKER", "WINNER"),
          series %in% c(5:10) 
         ) |>
  select(series, episode, baker, result) %>%
    arrange(series, episode)
```


```



Import, clean, tidy, and organize the viewership data in viewers.csv. Show the first 10 rows of this dataset. What was the average viewership in Season 1? In Season 5?




